---
title: "Example Use"
output: rmarkdown::html_vignette
descriptions: >
  How do I use codedtitles in R? This vignette will go       through some of the functionality of this package and      scenarios where it may be helpful.
vignette: >
  %\VignetteIndexEntry{Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

How do I use codedtitles in R? This vignette will go through some of the functionality of this package and scenarios where it may be helpful.

Load the package:

```{r setup}
library(codedtitles)
```

### Overview

-   "codedtitles" is a package designed to make the pre-processing of variables simpler prior to moving forward with analysis.

-   The motivation for this package was born out of our own clinical research experience. For example, a large surgical database with over 200 variables was provided for analysis. None of the column titles had proper coding and many contained special characters and spaces, which would make it syntactically challenging to deal with in R. Given that many of these databases are collected by clinicians or people without statistical software experience, they may not be familiar with the role of variable coding. As such, we have attempted to create a novel, simple solution to speed up this process.

### Example dataframe

```{r}
data <- df
colnames(data)
```

-   Here is an example dataframe that contains some example variable names that may be handed to a statistician. As we can see, a number of this variables have long names, some with special characters, some have very similar names too.

### Case 1: Base function

-   First, we will demonstrate the base function of the package. This will organize all the column names by shortening the stem words, making them lower case and removing special characters.

```{r}
new_names <- codevar(data)
```

-   The output here shows our new names which have been simplified. If we want to reference these new names to the old names, we can call "coderef".

```{r}
print(coderef)
```

### Case 2: Reducing length of characters for each variable

-   By default, the function allows for a max_length of the new names to be 15 characters. However, if the user desires shorter variable names, this can be manually changed in the arguments. To maintain the meaning of the variable, the function truncates the individual words separately. Let try an example with max_length at 6

```{r}
new_names <- codevar(data, max_length = 6)
```

-   The function attempts to find simpler version of each word in the name and shorten them to meet the max_length argument.

### Case 3: Removing word splitting

-   Lets say we make max_length at 3 so we have very short variables to work with.

```{r}
new_names <- codevar(data, max_length = 3)

```

-   While this does shorten our variables significantly, the function still attempts to separate individual words within the variable name and then recombines after truncating with an underscore. If we want to remove these underscores, and instead shorten the whole variables we can set the split argument to false.

```{r}
new_names <- codevar(data, max_length = 3, split = FALSE)

```

-   Again, if we are unsure of the original name, we can always reference the "coderef" dataframe.

### Case 4: Repeat Variables

-   Lets look at our last example where our max_length was 3 and there is no splitting. This leads to equal new variable names for variables like "Attack_Rate", "Attack_Rate_Total" and "attacK".

-   The code anticipates this and will sequentially add a number tag to repeat variables:

```{r}
coderef[c(25,26,35),]

```

-   This prevents any two variables from having the exact same name new re-coded name.

### Case 5: Add tag

-   Lastly, if user wants to add a tag to variables after it is run through the function, an argument is provided. We also demonstrate how you can choose to only apply this function to select columns if desired.

```{r}
# We will add a tag "_bl" to columns 4-7 to specify them as baseline data
new_names <- codevar(data[,4:7], max_length = 3, split = FALSE, tag = "_bl")

```
